The Ky Fan Norms and Beyond: Dual Norms and Combinations for Matrix Optimization

### 2025-12-01 article_final/main.tex
Resized the table

 \begin{table}[ht]
 \floatconts
   {tbl:matrix_methods}
   {\caption{Comparison of methods for computing rank-$k$ updates on dense random matrices (CPU, double precision). Lower is better in all columns.}}
   {%
     \begin{tabular}{cccccccc}
       \arrayrulecolor{black}\toprule
       Matrix sizes & $k$  & Method           & Time (s) & Matvecs & Iterations & $err_1$         & $err_2$       \\
       \arrayrulecolor{black}\hline
       $500 \times 500$      & 5  & Power Iterations           & 0.062   & 2005    & 200        & $9.2 \cdot 10^{-3}$ & $9.1 \cdot 10^{-3}$\\
       $500 \times 500$      & 5  & RSVD             & 0.017   & 1170    & 38         & $9.8 \cdot 10^{-3}$ & $9.6 \cdot 10^{-3}$\\
       $500 \times 500$      & 5  & TRLan            & 0.018   & 131     & 65         & $9.6 \cdot 10^{-5}$ & $9.4 \cdot 10^{-5}$\\
       \hdashline
       $500 \times 500$      & 50 & Power Iterations           & 0.44    & 43750   & 437      & $9.9 \cdot 10^{-3}$ & $9.0 \cdot 10^{-3}$\\
       $500 \times 500$      & 50 & RSVD             & 0.61    & 6120    & 50      & $9.9 \cdot 10^{-3}$ & $9.1 \cdot 10^{-3}$\\
       $500 \times 500$      & 50 & TRLan            & 0.16    & 462     & 231      & $3.3 \cdot 10^{-7}$ & $3.0 \cdot 10^{-7}$\\
       \hdashline
       $5000 \times 5000$    & 5  & Power Iterations           & 9.6     & 9065    & 906      & $8.6 \cdot 10^{-3}$ & $8.6 \cdot 10^{-3}$\\
       $5000 \times 5000$    & 5  & RSVD             & 2.1     & 5640    & 187      & $9.7 \cdot 10^{-3}$ & $9.7 \cdot 10^{-3}$\\
       $5000 \times 5000$    & 5  & TRLan            & 0.70    & 205     & 102      & $7.7 \cdot 10^{-3}$ & $7.7 \cdot 10^{-3}$\\
       \arrayrulecolor{black}\bottomrule
     \end{tabular}%
   }
 \end{table}

@@

 We generalized several successful algorithms, including Muon, Dion, and $\nu$SAM, to LMO-based algorithms using the family of norms dual to Ky Fan $k$-norms, yielding the Fanion family with low-rank updates. We further proposed the technique of regularizing these updates by combining them with NSGD or SignSGD, creating the F-Fanion and S-Fanion families through the $\norm{\cdot}_{\mathrm{F-KF-k}}^\dagger$ and $\norm{\cdot}_{\mathrm{C^\dagger-KF-k}}^\dagger$ norms. Our experiments demonstrate that F-Muon and S-Muon achieve competitive performance with Muon on CIFAR-10 airbench and NanoGPT tasks, confirming that the underlying norm constraint can be significantly modified without sacrificing effectiveness.

 However, our results also reveal that not every LMO-based algorithm is effective: Neon (rank-one Fanion) underperforms despite sharing the same theoretical convergence asymptotics as Muon in existing bounds. We suggest that future work on non-Euclidean LMO algorithms should explain in the corollaries to their convergence theorems the empirical superiority of Muon over other Fanions, and ideally account for the robustness of F-Muon and S-Muon as well. Without such refinements, it is difficult to believe that current theoretical bounds are relevant for practitioners.

>>>>

 \begin{table}[ht]
 \floatconts
   {tbl:matrix_methods}
   {\caption{Comparison of methods for computing rank-$k$ updates on dense random matrices (CPU, double precision). Lower is better in all columns.}}
   {\resizebox{\textwidth}{!}{%
     \begin{tabular}{cccccccc}
       \arrayrulecolor{black}\toprule
       Matrix sizes & $k$  & Method           & Time (s) & Matvecs & Iterations & $err_1$         & $err_2$       \\
       \arrayrulecolor{black}\hline
       $500 \times 500$      & 5  & Power Iterations           & 0.062   & 2005    & 200        & $9.2 \cdot 10^{-3}$ & $9.1 \cdot 10^{-3}$\\
       $500 \times 500$      & 5  & RSVD             & 0.017   & 1170    & 38         & $9.8 \cdot 10^{-3}$ & $9.6 \cdot 10^{-3}$\\
       $500 \times 500$      & 5  & TRLan            & 0.018   & 131     & 65         & $9.6 \cdot 10^{-5}$ & $9.4 \cdot 10^{-5}$\\
       \hdashline
       $500 \times 500$      & 50 & Power Iterations           & 0.44    & 43750   & 437      & $9.9 \cdot 10^{-3}$ & $9.0 \cdot 10^{-3}$\\
       $500 \times 500$      & 50 & RSVD             & 0.61    & 6120    & 50      & $9.9 \cdot 10^{-3}$ & $9.1 \cdot 10^{-3}$\\
       $500 \times 500$      & 50 & TRLan            & 0.16    & 462     & 231      & $3.3 \cdot 10^{-7}$ & $3.0 \cdot 10^{-7}$\\
       \hdashline
       $5000 \times 5000$    & 5  & Power Iterations           & 9.6     & 9065    & 906      & $8.6 \cdot 10^{-3}$ & $8.6 \cdot 10^{-3}$\\
       $5000 \times 5000$    & 5  & RSVD             & 2.1     & 5640    & 187      & $9.7 \cdot 10^{-3}$ & $9.7 \cdot 10^{-3}$\\
       $5000 \times 5000$    & 5  & TRLan            & 0.70    & 205     & 102      & $7.7 \cdot 10^{-3}$ & $7.7 \cdot 10^{-3}$\\
       \arrayrulecolor{black}\bottomrule
     \end{tabular}%
   }
   }
 \end{table}

@@

 We generalized several successful algorithms, including Muon, Dion, and $\nu$SAM, to LMO-based algorithms using the family of norms dual to Ky Fan $k$-norms, yielding the Fanion family with low-rank updates. We further proposed the technique of combining them with NSGD or SignSGD, creating the F-Fanion and S-Fanion families through the $\norm{\cdot}_{\mathrm{F-KF-k}}^\dagger$ and $\norm{\cdot}_{\mathrm{C^\dagger-KF-k}}^\dagger$ norms. Our experiments demonstrate that F-Muon and S-Muon achieve competitive performance with Muon on CIFAR-10 airbench and NanoGPT tasks, confirming that the underlying norm constraint can be significantly modified without sacrificing effectiveness.

 However, our results also reveal that not every LMO-based algorithm is effective: Neon (rank-one Fanion) underperforms despite sharing the same theoretical convergence asymptotics as Muon in existing bounds. We suggest that future work on non-Euclidean LMO algorithms should explain in the corollaries to their convergence theorems the observed empirical superiority of Muon over other Fanions, and ideally account for the robustness of F-Muon and S-Muon as well.
